\section*{To Do}\label{sec:todo}

\begin{enumerate}

	\item \textbf{Reach out to Armstrong and/or Kolesár.}
	Connections: Christoph Rothe (supervisor) has worked with Kolesár; PGP (supervisor) has worked with Armstrong; Soonwoo Kwon (coauthor) has worked with Kolesár.
	Show them the coupling-agnosticism result (Proposition + corollaries) and the connection to their modulus-of-continuity framework.
	Ask whether the modulus computation for anisotropic constraint sets is tractable, and explore cooperation.

	\item \textbf{Show that AK 2018 Appendix~B.1's $\mathcal{H}_y$ class is implicitly pooled/isotropic.}
	The pooled regression function $\bar{m}(x) = \int m(x,z) f(z|x)\, dz$ has its derivative bounded by a single scalar $C_y$, which is fungible across the direct channel ($\partial m / \partial x$) and the compositional channel (through $\partial f / \partial x$).
	Show formally that the anisotropic parameterization (separate $M_x$, $M_z$ plus density smoothness) yields a strictly smaller modulus $\omega(\delta; \mathcal{F}_{\mathrm{aniso}}) < \omega(\delta; \mathcal{H}_y)$, and hence strictly shorter CIs.
	This is the central applied result connecting the Proposition to the AK framework.

	\item \textbf{Compute the modulus of continuity for $\mathcal{F}(\mathcal{C}_{\mathrm{rect}})$ and $\mathcal{F}(\mathcal{C}_{\mathrm{disk}})$.}
	The ratio $\omega(\delta; \mathcal{F}_{\mathrm{iso}}) / \omega(\delta; \mathcal{F}_{\mathrm{aniso}})$ gives the quantitative gain $\rho(M_x, M_z)$.
	Three routes: (1) direct optimization over function spaces, (2) support function representation (most promising), (3) Gaussian white noise equivalence via \textcite{bertin_exact_2004}.

	\item \textbf{Read \textcite{bertin_exact_2004} carefully.}
	Exact asymptotic minimax constants for anisotropic Hölder classes.
	Extract the formula for $\mathfrak{R}^*_{\mathrm{aniso}}(M_x, M_z)$.

	\item \textbf{Derive multivariate bias formula and optimal weights.}
	Generalize the univariate RDD formulas of \textcite{armstrong_optimal_2018} (eq.~11: worst-case bias; eq.~14: local polynomial weights) to the bivariate case $f(x, z)$ under the anisotropic constraint set.

	\item \textbf{AMSE comparison on common parameter space.}
	Construct $\mathcal{F}^R(M_x, M_z, L)$ with anisotropic bounds on $m$ plus density smoothness constraint.
	Show $\overline{\mathrm{AMSE}}(\hat\tau_{\mathrm{seg}}) \leq \overline{\mathrm{AMSE}}(\hat\tau_{\mathrm{base}})$ with strict inequality when the compositional channel is active.

	\item \textbf{Implement optimized RDD with rectangular constraint.}
	Feed the anisotropic constraint set into the IK/AK weight optimization.
	Demonstrate CI improvement in simulation or empirical application.

	\item \textbf{Verify novelty against remaining references.}
	Korostelev and Korostelev (2011), Giné and Nickl (2016), Wasserman (2006).
	Confirm that the coupling-agnosticism characterization and the ``two types of ignorance'' distinction are absent.

	\item \textbf{Verify bib entry details.}
	Check page/volume numbers for Lepski (1991), LMS (1997), KLP (2001).

\end{enumerate}
